---
title: "Tidy Tuesday Meetup"
subtitle: "2023 week ?: London Marathon"
author: "Alice Walsh for R-Ladies Philly"
date: "2023-04-25"
output: 
  html_document:
    theme: lumen
    toc: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages, echo=FALSE}
# This code chunk won't be displayed in final product because set echo=FALSE
library(dplyr) # for data manipulation
library(ggplot2) # for plots
library(corrplot) # for correlation plot

theme_set(theme_light())
```

# Overview

Let's look at some fun data as part of a tidy tuesday event. This week, the 
data is from the London Marathon.

https://github.com/rfordatascience/tidytuesday/tree/master/data/2023/2023-04-25

https://nrennie.rbind.io/blog/web-scraping-rvest-london-marathon/

# Load the data

```{r}
winners <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-04-25/winners.csv') %>% 
        janitor::clean_names()
london_marathon <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-04-25/london_marathon.csv') %>% 
        janitor::clean_names()

glimpse(winners)
glimpse(london_marathon)
```

# Ideas

+ Patterns over time?
+ Do some simple tidymodels? (winners is only 163 rows, and no losers)
        + Could we predict the winning time based on the weather?
+ Add in other data to look at the weather in london?
  + https://www.weather.gov/documentation/services-web-api
  + https://www.kaggle.com/datasets/emmanuelfwerr/london-weather-data
+ look for outliers? really high or low values?
+ Look at other marathon data with rvest (other cities)

# Analysis

We added a dataset with the london weather that was sourced from a kaggle dataset:

https://www.kaggle.com/datasets/emmanuelfwerr/london-weather-data?resource=download

```{r}
weather <- readr::read_csv(
        "https://raw.githubusercontent.com/rjake/one-off-projects/main/R/tidy-tuesday/2023-04-25-London-Marathon/weather.csv")
head(weather)

# Merge with weather
marathon_weather <- inner_join(london_marathon, weather, by = "date")
winner_weather <- inner_join(london_marathon %>% select(date, year), 
                             weather, 
                             by = "date") %>% 
        left_join(winners, by = "year")

# merging it all!
prep_df <- 
        winners %>%
        distinct(year, category, time) %>%
        mutate(time = as.numeric(time)) %>% 
        tidyr::pivot_wider(
                names_from = category, 
                values_from = time
        ) %>%
        inner_join(weather %>% 
                           mutate(year = lubridate::year(date)),
                   by = "year") %>%
        inner_join(london_marathon) %>%
        mutate(month = lubridate::month(date)) %>% 
        janitor::clean_names() %>% 
        print()
```

## Overall correlation of the variables

Wow! A lot of these are highly correlated!

+ year has a large effect - times are getting better (lower) over time
+ year is also highly associated with all the participant stats
+ the weather variables are correlated with each other
+ we dont see a large correlation between the weather variables and the participant stats or the winning times

We did not have time to try {ggpairs}

```{r}
all_correlations <- cor(
        prep_df %>% 
                select(-raised, -snow_depth) %>% 
                select(where(is.numeric)) %>% 
                tidyr::drop_na(),
        method = "spearman")
corrplot(all_correlations)
```

## Impact of weather on the number of finishers?
```{r}
marathon_weather %>% 
        ggplot(aes(x = precipitation, y = finishers)) + 
        geom_point(colour = "#88398A") + 
        scale_x_log10() + 
        theme_light()

marathon_weather %>% 
        ggplot(aes(x = max_temp, y = finishers)) + 
        geom_point(colour = "#88398A") + 
        scale_x_log10() + 
        theme_light()
```


## Plots of winners data

Let's explore the trends in the winner's times

```{r}
# facet grid of winners by Category over Year
ggplot(winner_weather, aes(precipitation, time)) + 
  geom_point(colour = "#88398A") + 
  facet_grid(cols = vars(category)) + 
  theme_bw()

ggplot(winner_weather, aes(max_temp, time)) + 
  geom_point(colour = "#88398A") + 
  facet_grid(cols = vars(category)) + 
  theme_bw()
```

Now we had the idea to look at models including the year and weather.

```{r}
# more rain = fewer finishers
lm(finishers ~ precipitation, data = prep_df) %>% 
        broom::tidy()
lm(finishers ~ precipitation + year, data = prep_df) %>% 
        broom::tidy()

lm(finishers ~ precipitation > 0, data = prep_df) %>% 
        broom::tidy()
lm(finishers ~ (precipitation > 0) + year, data = prep_df) %>% 
        broom::tidy()

# higher temp = more finishers, but decreased effect when including year!
lm(finishers ~ max_temp, data = prep_df) %>% 
        broom::tidy()
lm(finishers ~ max_temp + year, data = prep_df) %>% 
        broom::tidy()
```

We also looked at the wheelchair_men times.

If we look back at the plot of precipitation versus the time, we see that there is 
one year with really high precipitation and also really high winning times. 
That appears to be an early year (1983) and the wheelchair times are slower in those
early years.

```{r}
# higher temp = faster time
lm(wheelchair_men ~ max_temp, data = prep_df) %>% 
        broom::tidy()
lm(wheelchair_men ~ max_temp + year, data = prep_df) %>% 
        broom::tidy()

# more rain = slower time?
lm(wheelchair_men ~ precipitation > 0, data = prep_df) %>% 
        broom::tidy()
lm(wheelchair_men ~ (precipitation > 0) + year, data = prep_df) %>% 
        broom::tidy()
lm(men ~ precipitation > 0, data = prep_df) %>% 
        broom::tidy()
lm(men ~ (precipitation > 0) + year, data = prep_df) %>% 
        broom::tidy()
```

```{r}
weather %>%
        ggplot(aes(x = date, y = max_temp)) + 
        geom_point() + 
        labs(title = "temperatures rose over time")
```

## Time differences by sex categories over time

```{r}
prep_df %>%
  mutate(
    diff_sex = women - men,
    diff_sex_wc = wheelchair_women - wheelchair_men
  ) %>%
  ggplot(aes(year, diff_sex)) +
        geom_smooth(method = "lm", se = FALSE) + 
  geom_point(aes(color = max_temp))
```

## Exploration of the winners data
```{r}
# trend of winners by Category over Year
ggplot(winners, aes(x=year)) + 
  geom_line(aes(y=time, col=category)) + 
  theme_bw()

# facet grid of winners by Category over Year
ggplot(winners, aes(year, time)) + 
  geom_point(colour = "#88398A") + 
  facet_grid(cols = vars(category)) + 
  theme_bw()

# facet grid of winners by Category and Gender over Year
winners %>% 
  mutate(Gender = gsub('Wheelchair ','',Category),
         Wheelchair = ifelse(nchar(Category) > 6,1,0)) %>% 
  ggplot(aes(Year, Time)) + 
  geom_point(colour = "#88398A") + 
  facet_grid(cols = vars(Gender), rows = vars(Wheelchair), scales = "free_y") + 
  theme_bw()
```


## Dealing with the y-axis?

We noticed that the winning times were formated weird sometimes. 
Here, they are converted to hours!

```{r}
winners %>%
  rename_all(tolower) %>%
  ggplot(aes(x = year, y = time)) +
  geom_line(aes(group = category)) +
  facet_wrap(~category) +
  scale_y_time(
    labels = ~paste0(lubridate::hour(.x), "H")
  )
```

# Contact information
For any questions, contact <philly@rladies.org>
