---
title: "Do more with dplyr::slice()"
author: "June Choe<br>@yjunechoe"
execute:
  echo: true
format:
  revealjs:
    logo: "https://www.rladiesphilly.org/author/r-ladies-philly/avatar_hue4e99079bc6317a6be3355d116429734_93925_270x270_fill_q90_lanczos_center.jpg"
    code-line-numbers: false
    slide-number: true
---

```{r}
#| include: false
library(dplyr)
library(ggplot2)
library(forcats)
```

## {background-color="#6C4176"}

<style>
span.center {
  font-size: 2em;
  position: absolute;
  top: 50%;
  left: 50%;
  transform: translate(calc(-50%), calc(-50%));
}
</style>

[Basics of `dplyr::slice()`]{.center}

## Basic usage: 1) row selection

```{r}
mtcars %>% 
  slice(1:6) # like `head()`, selects rows [1,...,6]
```

. . .

```{r}
mtcars %>% 
  slice( n() ) # selects last row [32]
```

. . .

```{r}
mtcars %>% 
  slice( (n() - 2) : n() ) # selects rows [30,31,32]
```

## Basic usage: 2) row reordering

```{r}
mtcars %>% 
  slice( c(1,3,5,2,4,6) )
```

. . .

```{r}
mtcars %>% 
  slice(1, 3, 5, 2, 4, 6) # FYI: "spliced" version works the same!
```

## Basic usage: 3) misc. behavior

```{r}
mtcars %>% 
  slice( -(3:n()) ) # same as `1:2` (all except third row to last row)
```

. . .

```{r}
mtcars %>% 
  slice( c(1, 1) ) # the first row, twice
```

. . . 

```{r}
mtcars %>% 
  slice( n() + 1 ) # the 33rd row (doesn't exist; doesn't error!)
```

```{r}
mtcars %>% 
  slice( c(1, n() + 1) ) # non-existent rows are just ignored
```

## {background-color="#6C4176"}

[`dplyr::slice()`, with a twist!]{.center}

## Hidden usage: row filtering

We know how `filter()` works...

<br>

```{r}
mtcars$mpg == 22.8 # logical vector, same length as data
```

<br>

```{r}
mtcars %>% 
  filter(mpg == 22.8) # subsets rows whose value is TRUE
```

## Hidden usage: row filtering

You can expect a similar behavior from `slice()`!

<br>

```{r}
which( mtcars$mpg == 22.8 )
```

<br>

. . .

```{r}
mtcars %>% 
  slice( which( mpg == 22.8 ) ) # evaluates to `slice( c(3, 9) )`
```

## Two sides of the same coin!

:::: {.columns}

::: {.column width="48%"}

Filter

```{r}
#| eval: false
df %>% 
  filter( ... )
```

<br>

```{r}
#| eval: false
df %>% 
  filter( A, B ) # `A & B`
```

<br>

```{r}
#| eval: false
df %>% 
  filter( A | B )
```

:::

::: {.column width="4%"}

:::

::: {.column width="48%"}

Slice

```{r}
#| eval: false
df %>% 
  slice( which( ... ) )
```

<br>

```{r}
#| eval: false
df %>% 
  slice( which( A & B ) )
```

<br>

```{r}
#| eval: false
df %>% 
  slice( which( A | B ) )
```

:::

::::

. . .

<br>

```{r}
#| eval: false
df %>% filter( row_number() %in% which( ... ) ) # round-about
```

## But why?

Filtering with slice has two interesting properties:

. . .

1) Vectors of filtered elements (=> set operations)

. . .

2) Elements as indices (=> integer operations)

. . .

Opens up new workflows

- **row-relational operations**

- **row shuffling**

## {background-color="#6C4176"}

[Row-relational operations]{.center}

## But why?

If you have data that encodes information in _row order_ (e.g., time), you can compute **relationships between rows**

```{r}
#| echo: false
flights_df <- nycflights13::flights %>% 
  filter(month == 6, day == 8, between(dep_time, 600, 1200),  origin == "JFK") %>% 
  select(dep_time, flight, carrier)
```

```{r}
flights_df
```


## Row-relational operations (Q1)

Which flights departed _right after_ American Airlines flights?

. . .

```{r}
flights_df$carrier == "AA"
```

. . .

```{r}
which( flights_df$carrier == "AA" )
```

. . .

```{r}
which( flights_df$carrier == "AA" ) + 1
```

## Row-relational operations (Q1)

Which flights departed _right after_ American Airlines flights?

```{r}
flights_df %>% 
  slice(
    # Next row after where `carrier == "AA"`
  )
```

## Row-relational operations (Q1)

Which flights departed _right after_ American Airlines flights?

```{r}
flights_df %>% 
  slice(
    # Next row after where `carrier == "AA"`
    which( carrier == "AA" ) + 1
  )
```


## Row-relational operations (Q2)

Do (1), but also include the American Airlines flights:

```{r}
#| eval: false
flights_df %>% 
  slice(
    # Previous solution to (1)
    which( carrier == "AA" ) + 1
  )
```

## Row-relational operations (Q2)

Do (1), but also include the American Airlines flights:

```{r}
flights_df %>% 
  slice(
    # Also including AA flights
    which( carrier == "AA" ),
    which( carrier == "AA" ) + 1
  )
```

## Row-relational operations (Q3)

Do (2), but preserve original row order:

```{r}
flights_df %>% 
  slice(
    # Previous solution to (2)
    which( carrier == "AA" ),
    which( carrier == "AA" ) + 1
  )
```


## Row-relational operations (Q3)

Do (2), but preserve original row order:

```{r}
flights_df %>% 
  slice(
    # First, concatenate the indices
    c( which( carrier == "AA" ),
       which( carrier == "AA" ) + 1 )
  )
```

## Row-relational operations (Q3)

Do (2), but preserve original row order:

```{r}
flights_df %>% 
  slice(
    # Then, sort
    sort( c( which( carrier == "AA" ),
             which( carrier == "AA" ) + 1 ) )
  )
```

## Row-relational operations (Q3)

Do (2), but preserve original row order:

```{r}
flights_df %>% 
  slice(
    # Optionally, drop duplicates
    unique( sort( c( which( carrier == "AA" ),
                     which( carrier == "AA" ) + 1 ) ) )
  )
```


## Row-relational operations (Q4)

Do (3), but with less code!

. . .

**Trick**: hoist them into a matrix and collapse item-wise!

. . .

```{r}
matrix(
  c( which( flights_df$carrier == "AA" ),
     which( flights_df$carrier == "AA" ) + 1 ),
  nrow = 2, byrow = TRUE
)
```

. . .

This essentially "stacks" two vectors:

```{r}
#| echo: false
m <- matrix(
  c( which( flights_df$carrier == "AA" ),
     which( flights_df$carrier == "AA" ) + 1 ),
  nrow = 2, byrow = TRUE
)
rownames(m) <- c("AA", "AA+1")
m
```


## Row-relational operations (Q4)

Do (3), but with less code!

**Trick**: hoist them into a matrix and collapse item-wise!

```{r, out.height='200px'}
#| code-line-numbers: "6"
matrix(
  c( which( flights_df$carrier == "AA" ),
     which( flights_df$carrier == "AA" ) + 1 ),
  nrow = 2, byrow = TRUE
) %>% 
  as.vector() # default column-wise collapsing
```

## Row-relational operations (Q4)

Do (3), but with less code!

**Trick**: _hoist_ them into a matrix and collapse item-wise!

**Trick 2**: use `base::outer()`!

```{r}
outer(
  X = c(0, 1),                                           # row
  Y = which( flights_df$carrier == "AA" ),               # col
  FUN = function(row, col) { row + col }                 # cell
)
```

## Row-relational operations (Q4)

Do (3), but with less code!

**Trick**: hoist them into a matrix and _collapse_ item-wise!

**Trick 2**: use `base::outer()`!

```{r}
#| code-line-numbers: "6"
outer(
  X = c(0, 1),                                           # row
  Y = which( flights_df$carrier == "AA" ),               # col
  FUN = function(row, col) { row + col }                 # cell
) %>% 
  as.vector()
```

## Row-relational operations (Q4)

Do (3), but with less code!

```{r}
flights_df %>% 
  slice(
    # Previous solution to (3)
    sort( c( which( carrier == "AA" ),
             which( carrier == "AA" ) + 1 ) )
  )
```

## Row-relational operations (Q4)

Do (3), but with less code!

```{r}
flights_df %>% 
  slice(
    # With `outer()`
    as.vector(outer( 0:1, which(carrier == "AA"), `+` ))
  )
```


## Row-relational operations (case)

Let's try a full row-relational workflow example

```{r}
#| echo: false
gapminder_recent <- gapminder::gapminder %>% 
  left_join(gapminder::country_codes, by = "country") %>% 
  filter(year > 1990) %>% 
  select(country, country_code = iso_alpha, year, gdpPercap)
```

```{r}
gapminder_recent
```

## Row-relational operations (case)

A bar plot of Germany's GDP growth is straightforward ...

```{r}
#| echo: false
gapminder_recent %>% 
  group_by(year) %>% 
  arrange(year, gdpPercap, by_group = TRUE) %>% # `year`-sorting optional
  slice( as.vector(outer(-1:1, which(country == "Germany"), `+`)) ) %>% 
  mutate(grp = forcats::fct_inorder(c("lo", "is", "hi"))) %>% 
  ungroup() %>% 
  ggplot(aes(as.factor(year), gdpPercap, group = grp)) +
  geom_col(aes(fill = grp == "is"), position = position_dodge()) +
  geom_text(
    aes(label = country_code, alpha = as.integer(grp == "is")), vjust = 1.3,
    position = position_dodge(width = .9)
  ) + scale_alpha_identity() +
  scale_fill_manual(
    values = c("NA", "steelblue"),
    guide = guide_none()
  ) +
  theme_classic() +
  labs(title = "Germany's GDP growth", x = "Year", y = "GDP per capita")
```

## Row-relational operations (case)

... but how can we also plot its GDP-neighbors?

```{r gapminder-plot}
#| echo: false
gapminder_recent %>% 
  group_by(year) %>% 
  arrange(year, gdpPercap, by_group = TRUE) %>% # `year`-sorting optional
  slice( as.vector(outer(-1:1, which(country == "Germany"), `+`)) ) %>% 
  mutate(grp = forcats::fct_inorder(c("lo", "is", "hi"))) %>% 
  ungroup() %>% 
  ggplot(aes(as.factor(year), gdpPercap, group = grp)) +
  geom_col(aes(fill = grp == "is"), position = position_dodge()) +
  geom_text(
    aes(label = country_code), vjust = 1.3,
    position = position_dodge(width = .9)
  ) +
  scale_fill_manual(
    values = c("grey75", "steelblue"),
    guide = guide_none()
  ) +
  theme_classic() +
  labs(title = "Germany's GDP growth", x = "Year", y = "GDP per capita")
```


## Row-relational operations (case)

```{r}
#| code-line-numbers: "1"
gapminder_recent
```

## Row-relational operations (case)

```{r}
#| code-line-numbers: "2-3"
gapminder_recent %>% 
  group_by(year) %>% 
  arrange(year, gdpPercap, by_group = TRUE) # `year`-sorting optional
```

## Row-relational operations (case)

```{r}
#| code-line-numbers: "2-3"
gapminder_recent %>% 
  group_by(year) %>% 
  arrange(gdpPercap, by_group = TRUE) # same within-`year` ordering
```

## Row-relational operations (case)

```{r}
#| code-line-numbers: "4"
gapminder_recent %>% 
  group_by(year) %>% 
  arrange(year, gdpPercap, by_group = TRUE) %>% # `year`-sorting optional
  slice( as.vector(outer( -1:1, which(country == "Germany"), `+` )) )
```

## Row-relational operations (case)

```{r}
#| code-line-numbers: "5-6"
gapminder_recent %>% 
  group_by(year) %>% 
  arrange(year, gdpPercap, by_group = TRUE) %>% # `year`-sorting optional
  slice( as.vector(outer(-1:1, which(country == "Germany"), `+`)) ) %>% 
  mutate(grp = forcats::fct_inorder(c("lo", "is", "hi"))) %>% 
  ungroup()
```

## Row-relational operations (case)

```{r}
#| code-line-numbers: "7-8"
gapminder_recent %>% 
  group_by(year) %>% 
  arrange(year, gdpPercap, by_group = TRUE) %>% # `year`-sorting optional
  slice( as.vector(outer(-1:1, which(country == "Germany"), `+`)) ) %>% 
  mutate(grp = forcats::fct_inorder(c("lo", "is", "hi"))) %>% 
  ungroup() %>% 
  ggplot(aes(as.factor(year), gdpPercap, group = grp)) +
  geom_col(aes(fill = grp == "is"), position = position_dodge())
```

## Row-relational operations (case)

```{r}
#| eval: false
#| code-line-numbers: "9-18"
<<gapminder-plot>>
```

## Row-relational operations (case)

```{r}
#| echo: false
<<gapminder-plot>>
```

## {background-color="#6C4176"}

[Row-relational operations]{.center}

## Row shuffling - the problem

Input:

```{r}
fruits <- c("ðŸŽ", "ðŸ‹", "ðŸ‡")[c(2,1,3,3,2,3,1,2,2,1,2,2,3,3,3)]
table(fruits)
cat(fruits)
```

. . .

Output:

```{r}
fruits_out <- c("ðŸŽ", "ðŸ‹", "ðŸ‡")[c(3,2,1,3,2,3,2,1,3,2,3,2,1,3,2)]
cat(fruits_out)
```

## Row shuffling - interweaving vectors

```{r}
#| eval: false
grapes <- which(fruits == "ðŸ‡")
grapes
```

```{r}
#| echo: false
grapes <- which(fruits == "ðŸ‡")
setNames(grapes, rep("ðŸ‡", 6))
```

. . .

```{r}
#| eval: false
lemons <- which(fruits == "ðŸ‹")
lemons
```

```{r}
#| echo: false
lemons <- which(fruits == "ðŸ‹")
setNames(lemons, rep("ðŸ‹", 6))
```

. . .

```{r}
#| eval: false
apples <- which(fruits == "ðŸŽ")
apples
```

```{r}
#| echo: false
apples <- which(fruits == "ðŸŽ")
setNames(apples, rep("ðŸŽ", 3))
```

. . .

```{r}
#| eval: false
apples <- c(apples[1], 0, apples[2], 0, apples[3], 0)
apples
```

```{r}
#| echo: false
apples <- c(apples[1], 0, apples[2], 0, apples[3], 0)
setNames(apples, c("ðŸŽ", "", "ðŸŽ", "", "ðŸŽ", ""))
```

## Row shuffling - `matrix()`, again!

```{r}
fruits_matrix <- matrix(c(grapes, lemons, apples), nrow = 3, byrow = TRUE)
fruits_matrix
```

. . .

```{r}
#| echo: false
rownames(fruits_matrix) <- names(table(fruits))
fruits_matrix
```

. . .

```{r}
#| echo: false
matrix(c("", fruits)[t(fruits_matrix+1)], nrow = 3, byrow = TRUE)
```


## Row shuffling - `matrix()`, again!

```{r}
#| echo: false
fruits_matrix <- matrix(c(grapes, lemons, apples), nrow = 3, byrow = TRUE)
```

```{r}
fruits_matrix
```

. . .

```{r}
as.vector(fruits_matrix)
```

<br>

. . .

```{r}
cat( fruits[as.vector(fruits_matrix)] )
cat( fruits_out )
```

## Row shuffling - same thing for rows

```{r}
#| echo: false
set.seed(1234)
fruits_df <- tibble(
  id = seq_along(fruits),
  fruit = c("ðŸŽ" = "Apple", "ðŸ‹" = "Lemon", "ðŸ‡" = "Grape")[fruits],
  quality = runif(15, min = .6, max = 1),
  price = c("Apple" = 1, "Lemon" = .5, "Grape" = 2)[fruit] * quality
)
```

```{r}
fruits_df
```

## Row shuffling - same thing for rows

```{r}
fruits_df %>% 
  slice( as.vector(fruits_matrix) )
```

## Row shuffling - same thing for rows

```{r}
fruits_df %>% 
  slice( penngradlings::rshuffle(fruit, mode = "squish") )
```

<span class="footnote">[Code for `rshuffle()`](https://github.com/yjunechoe/penngradlings/blob/master/R/data-wrangling-misc.R)</span>

## {background-color="#6C4176"}

[Thank you!]{.center}
