---
title: "Are home prices related to lead testing in Philadelphia?"
subtitle: "using open data"
author: "R-Ladies Philly"
date: "1/28/2018"
output: 
  html_document:
    theme: lumen
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages, echo=FALSE}
# This code chunk won't be displayed in final product because set echo=FALSE
library(ggplot2)
```

# Overview
This document is an example analysis using data found on www.opendataphilly.org
  

# Data sources
```{r data_sources, echo=FALSE}
lead_source <- "https://phl.carto.com/api/v2/sql?q=SELECT+*+FROM+child_blood_lead_levels_by_zip&filename=child_blood_lead_levels_by_zip&format=csv&skipfields=cartodb_id,the_geom,the_geom_webmercator"

zillow_source <- "http://files.zillowstatic.com/research/public/Zip/Zip_MedianListingPrice_AllHomes.csv"
```
## Lead testing data
Data about lead testing in children by zip code was loaded from `r lead_source`.  
More information on this data can be found at: <https://www.opendataphilly.org/dataset/philadelphia-child-blood-lead-levels>
  

```{r load_lead}
lead_data <- read.csv(lead_source)

knitr::kable(head(lead_data), caption = "First few rows of data")
```

The lead testing data consists of *`r nrow(lead_data)`* rows and *`r ncol(lead_data)`* columns.

## Home prices
Zillow makes home prices and other data available on their website: https://www.zillow.com/research/data/
  
We loaded the data on median listing prices for all Philadelphia zip codes from: `r zillow_source`
  
```{r load_prices}
zillow_data <- read.csv(zillow_source)

knitr::kable(head(zillow_data[,1:7]), caption = "First few rows and columns of data")
```
  
The data contains data from all over the country, we will only keep the Philadelphia data. The columns contain median home prices for a given year and month.

The two datasets were merged - this will only select rows with zip codes in common between the two datasets.
```{r merge}
full_dataset <- merge(lead_data, zillow_data, by.x = "zip_code", by.y = "RegionName")
```

The merged data consists of *`r nrow(full_dataset)`* rows and *`r ncol(full_dataset)`* columns.

# Visualization
```{r plots}

ggplot(full_dataset, aes(x = X2017.11, y = perc_5plus)) +
  geom_point() +
  ylab("Percent of children screened with BLLs ≥5 µg/dL") + 
  xlab("Median home price, November 2017") +
  theme_bw()

ggplot(full_dataset, aes(x = X2017.11, y = num_screen)) +
  geom_point(aes(size=num_screen)) +
  ylab("Number of children screened") + 
  xlab("Median home price, November 2017") +
  theme_bw()
  

```

It looks like lead testing and positive lead tests are higher in zip codes with lower home values.

# Future directions
The plots above are fairly boring. We could add some categories to facet by or color by.  
For example:  

  + Binning home prices to create categories
  + Adding in meta-data from other sources, such as the total population of each zip code 
  + Adding any factors that might differentiate zip codes with similar lead levels
  + Performing statistical tests
  
  
# Contact information
For any questions, contact <philly@rladies.org>
  

